#missing values, we used is.null function to show any null values in the dataset, we did not have any.

title: "data salaries" output: html_notebook: default github_document: default ---

## Goal of collecting this Dataset

Predicting Data Science job salaries using regression.

## The source of the dataset Link to the source Dataset:

<https://www.kaggle.com/datasets/arnabchaki/data-science-salaries-2023?resource=download>

## Informition about the Dataset :

Number of Attributes: 11

Type of Attributes: nominal, ordinal, numeric

Number of objects: 3755

Class name: salary in usd

| Attributes names   | Description                                                                                 | Data type        | Possible values                                                                                |
|------------------|------------------|------------------|------------------|
| work_year          | The year the salary was paid                                                                | ordinal          | Range between 2020-2023                                                                        |
| experience_level   | The experience level in the job during the year                                             | ordinal          | (SE:"Senior",MI:"Mid level",EN:"Entry level,EX:"Executive level")                              |
| employment_type    | The type of employment for the role                                                         | nominal          | (FT:"Full time",PT:"Part time",CT:"Contractual",FL:"Freelancer")                               |
| job_title          | The role worked in during the year                                                          | nominal          | (Data Engineer, Data Scientist,Data Analyst, Machine Learning Engineer, Analytics Engineer...) |
| salary             | The total gross salary amount paid                                                          | numeric interval | Range Between [6000 - 30.4m]                                                                   |
| salary_currency    | The currency of the salary paid as an ISO 4217 currency code                                | nominal          | USD,EUR,GBP,INR,CAD...)                                                                        |
| salaryinusd        | The salary in USD                                                                           | numeric interval | Between[5132-450k]                                                                             |
| employee_residence | Employee's primary country of residence in during the work year as an ISO 3166 country code | nominal          | (US,GB,CA,ES,IN...)                                                                            |
| remote_ratio       | The overall amount of work done remotely                                                    | numeric Ratio    | (0, 50, 100)                                                                                   |
| company_location   | The country of the employer's main office or contracting branch                             | nominal          | (US,GB,CA,ES,IN...)                                                                            |
| company_size       | The median number of people that worked for the company during the year                     | ordinal          | (S,M,L)                                                                                        |

### uploding the data and raw samples

raw data samples

```{r}
ds_salaries= read.csv(url("https://raw.githubusercontent.com/DeemAlshaye/Miningproject/main/dataset/ds_salaries.csv"), header=TRUE)
head(ds_salaries)
tail(ds_salaries)
```

```{r}
dim(ds_salaries)
```

### summray the data

we used summary function to show each numeric value min, median, max, Q1 and Q3

```{r}
summary(ds_salaries)
```

### statical mesurment

Statistical measurements allows to explore our data in a quantitative manner,it helps to understand its distribution, variability, and key characteristics.

```{r}
mean(ds_salaries$salary)
median(ds_salaries$salary)
quantile(ds_salaries$salary)
```

```{r}
mean(ds_salaries$salary_in_usd)
median(ds_salaries$salary_in_usd)
quantile(ds_salaries$salary_in_usd)
```

```{r}
mean(ds_salaries$remote_ratio)
median(ds_salaries$remote_ratio)
quantile(ds_salaries$remote_ratio)
```

```{r}
mean(ds_salaries$work_year)
median(ds_salaries$work_year)
quantile(ds_salaries$work_year)
```

```{r}
var(ds_salaries$salary)
var(ds_salaries$salary_in_usd)
var(ds_salaries$remote_ratio)
var(ds_salaries$work_year)
```

### changing the salary and salary in usd to be in 1000s

```{r}
ds_salaries$salary_1000s <- ds_salaries$salary /1000
head(ds_salaries$salary_1000s)

ds_salaries$salaryusd_1000s <- ds_salaries$salary_in_usd /1000
head(ds_salaries$salaryusd_1000s)
```

### graphs: we compared some variables with each other to find how they are distributed

#### boxplot:

```{r}
boxplot(ds_salaries$salary_1000s,ylim=c(0,1000),xlab="Salary",ylab="Salary in 1000s")
```

```{r}
boxplot(ds_salaries$salaryusd_1000s,ylim=c(0,1000),xlab="Salary in USD",ylab="Salary in usd 1000s")
```

```{r}
boxplot(ds_salaries$work_year,xlab="Work year",ylab="Frequency")
```

### histogram

```{r}
hist(ds_salaries$salary_1000s,ylim=c(0,100),ylab="Frequency",xlab="salary in 1000s",main = "Salary Frequency")
#the histogram represent the amout of salaries for each employee, it shows that most of employees has small amount of salary
```

### bar plot

```{r}
library(magrittr)

ds_salaries$remote_ratio %>% table() %>% barplot(xlab="remote ratio", main="barplot of remote ratio")
#the bar plot represent the total of remote ratio for each employee, it show that 0(onsite) is the most remote ratio

ds_salaries$work_year %>% table() %>% barplot(xlab="work year", ylab="number of employees", main="barplot of work year")
#the bar plot represent the work year and number of employee in each year, it show that 2020 year has the lowest number of employees

  library(ggplot2)
library(tidyverse)
top_5_job_salaries<-ds_salaries%>%
  group_by(job_title)%>%
  summarise(Avg_Sal=mean(salary_1000s))%>%
  arrange(desc(Avg_Sal))%>%
  head()
top_5_job_salaries
#we took the bar plot of 'Top 5 job Title Salaries' represent the job title and the salary for each job, it shows that Head of Machine Learning is the highest salary
```

### Scatter plot

```{r}
  with(ds_salaries,plot(ds_salaries$salary_1000s,ds_salaries$salaryusd_1000s,xlab="salary in 1000s",ylab = "salary in usd 1000s",main="Scatter plot with salary and salary in USD") )
```

### pie chart

```{r}
library(dplyr)
ds_salaries2 <- ds_salaries %>% sample_n(50)

tab <- ds_salaries2$company_size%>% table()
precentages <- tab %>% prop.table() %>% round(3)*100
txt <- paste0(names(tab),'\n',precentages,'%')
pie(tab,labels=txt)
#the pie chart represent the percentages for company size by taken sample of company, it shows that 2(Medium) has the highest frequency
```

### Visualisation

```{r}
ggplot(top_5_job_salaries, aes(x=job_title, y=Avg_Sal)) +
  geom_col() +
  labs(title="Top 5 Job Title Salaries", x="Job Title", y="Salary in 1000s") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r}
ggplot(ds_salaries, aes(x=experience_level, y=salaryusd_1000s)) +
  geom_boxplot(fill="pink") +
  labs(title="Distribution of Salary by Experience Level", x="Experience Level", y="Salary in 1000s") +
  theme_minimal()
```

### Average Salaries in each Year

```{r}
  library(dplyr)

yearly_salary_avg<-ds_salaries%>%
  group_by(work_year)%>%
  summarise(avg_salary=mean(salary))
yearly_salary_avg
    
```

## Data cleaning

### categorizeing job title

##Categorize job title coulmn

```{r}
library(tidyverse)
categorize_job_title <- function(title) {
  title <- tolower(title)
  if (grepl("data scientist|research scientist|researcher|scientist|science", title)) {
    return("Data Scientist")
  } else if (grepl("data engineer|etl|machine learning software engineer|ai|data architect|engineer|data modeler|machine learning engineer|devops", title)) {
    return("Data Engineer")
  } else if (grepl("data analyst|bi|analyst|head of data|data lead|data strategist|data analytics|specialist|data manager", title)) {
    return("Data Analyst")
  } else {
    return("Other")
  }
}
ds_salaries <- ds_salaries %>% 
  mutate(job_title = sapply(job_title, categorize_job_title))
head(ds_salaries)
```

Our column job title had 93 different categories, By using this categorization process, the code assigns each job title in the dataset to a general job title, we categorized the job title into four categorize ("Data Scientist","Data Engineer","Data Analyst",Other) , allowing for easier analysis and grouping of data based on their job title. \### categorizeing company location

```{r}


# Define the country categories
asia <- c("AE", "CN", "HK", "ID", "IN", "JP", "KR", "MY", "PH", "SG", "TH", "TW", "VN")
europe <- c("AL", "AM", "AT", "BA", "BE", "BG", "BY", "CH", "CZ", "DE", "DK", "EE", "ES", "FI", "FR", "GB", "GE", "GR", "HR", "HU", "IE", "IL", "IT", "LT", "LU", "LV", "MD", "MK", "MT", "NL", "NO", "PL", "PT", "RO", "RS", "RU", "SE", "SI", "SK", "TR", "UA", 'ES')

north_america <- c("CA", "US")
south_america <- c("AR", "BO", "BR", "BS", "CL", "CO", "CR", "DO", "EC", "GT", "HN", "JM", "MX", "NI", "PA", "PE", "PY", "SV", "UY", "VE")
oceania <- c("AS", "AU", "FJ", "GU", "KI", "MH", "MP", "NC", "NF", "NZ", "PG", "PW", "SB", "TO", "TV", "VU", "WS")
africa <- c("BF", "BI", "BJ", "BW", "CD", "CG", "CI", "CM", "CV", "DJ", "DZ", "EG", "ET", "GA", "GH", "GM", "GN", "GQ", "KE", "LR", "LS", "LY", "MA", "MG", "ML", "MR", "MU", "MW", "MZ", "NA", "NE", "NG", "RE", "RW", "SC", "SD", "SH", "SL", "SN", "SO", "SS", "ST", "SZ", "TD", "TG", "TN", "TZ", "UG", "ZA", "ZM", "ZW")

# Function to categorize the company locations
categorize_location <- function(location) {
  if (location %in% asia) {
    return("Asia")
  } else if (location %in% europe) {
    return("Europe")
  } else if (location %in% north_america) {
    return("North America")
  } else if (location %in% south_america) {
    return("South America")
  } else if (location %in% oceania) {
    return("Oceania")
  } else if (location %in% africa) {
    return("Africa")
  } else {
    return("Other")
  }
}

# Apply the categorization to the company location column
ds_salaries$company_location <- sapply(ds_salaries$company_location,categorize_location)

# Print the updated dataset with categories
tail(ds_salaries)
```

our column company location did have 72 different categories, By using this categorization process, the code assigns each company location in the dataset to a specific geographical region, such as Asia, Europe, North America, South America, Oceania, and Africa, allowing for easier analysis and grouping of data based on different regions of the world.

### finding missing data

we used is.null function to show any null values in the dataset.

```{r}
sum(is.na(ds_salaries))
```

we did not have any missing values.

### finding outliers:

By detecting outliers, analysts can identify potential errors or anomalies in the data collection process, data entry, or measurement inaccuracies

```{r}
boxplot.stats(ds_salaries$salary)$out
boxplot.stats(ds_salaries$salary_in_usd)$out              
boxplot.stats(ds_salaries$remote_ratio)$out
boxplot.stats(ds_salaries$work_year)$out
```

we did detect the outliers in our numeric attributes \### sum outliers:

```{r}
sum(boxplot.stats(ds_salaries$salary)$out)

sum(boxplot.stats(ds_salaries$salary_in_usd)$out)

sum(boxplot.stats(ds_salaries$remote_ratio)$out)

sum(boxplot.stats(ds_salaries$work_year)$out)
```

### removing the outliers:

Outliers can sometimes distort our analysis and models, so by removing them, we get a clearer picture of the data and can make better predictions or understand patterns more accurately

```{r}
library(outliers)
outliers <- boxplot(ds_salaries$salary, plot=FALSE)$out
ds_salaries <- ds_salaries[-which(ds_salaries$salary%in% outliers),]
boxplot.stats(ds_salaries$salary)$out

outliers <- boxplot(ds_salaries$salary, plot=FALSE)$out
ds_salaries <- ds_salaries[-which(ds_salaries$salary%in% outliers),]
boxplot.stats(ds_salaries$salary)$out

outliers <- boxplot(ds_salaries$salary_in_usd, plot=FALSE)$out
ds_salaries <- ds_salaries[-which(ds_salaries$salary_in_usd%in% outliers),]
boxplot.stats(ds_salaries$salary_in_usd)$out

outliers <- boxplot(ds_salaries$work_year, plot=FALSE)$out
ds_salaries <- ds_salaries[-which(ds_salaries$work_year%in% outliers),]
boxplot.stats(ds_salaries$work_year)$out
```

We did this to make sure our data is more accurate and representative of the majority of the values.

## Data transformation

### Normalization

Normalizing data helps in improving the performance of machine learning models,reducing the training time and improving efficiency.

```{r}
normalize <- function(x) {return((x-min(x))/(max(x)-min(x)))}
ds_salaries$salary<-normalize(ds_salaries$salary)
ds_salaries$salary_in_usd<-normalize(ds_salaries$salary_in_usd)

num1_cols=ds_salaries[, c(5,7 )] 
head(num1_cols)
```

we normalized the attributes salary, salary in usd, so it takes values between 0 and 1 ,which helps in handling the data

### Discretization

transforming continuous data into categorical values.

```{r}
breaks <- quantile(ds_salaries$salary_in_usd, 
                   probs = c(0, 1/7, 2/7, 3/7, 4/7, 5/7,6/7,1), 
                   na.rm = TRUE)

ds_salaries$salary_in_usd_disc <- cut(ds_salaries$salary_in_usd, 
                             breaks = breaks, 
                             include.lowest = TRUE, 
                             labels=c("low", "Low-mid", "mid", "mid-high", "high","very-high","top"))
head(ds_salaries$salary_in_usd_disc)
```

Discretization makes it easier to analyze and interpret the data \### Encoding onverting categorical or ordinal variables into a numerical representation that can be used in data analysis and modeling. This transformation is necessary because many machine learning algorithms and statistical techniques require numerical inputs.

```{r}
ds_salaries$company_size = factor(ds_salaries$company_size,levels = c("S","M","L"),labels = c(1,2,3))
ds_salaries$experience_level = factor(ds_salaries$experience_level,levels = c("EN","MI","SE","EX"),labels = c(1,2,3,4))

ds_salaries$job_title  <- factor(ds_salaries$job_title)
ds_salaries$employment_type  <- factor(ds_salaries$employment_type)
ds_salaries$employee_residence  <- factor(ds_salaries$employee_residence)
ds_salaries$company_location  <- factor(ds_salaries$company_location)
ds_salaries$salary_currency  <- factor(ds_salaries$salary_currency)
ds_salaries$job_title  <- factor(ds_salaries$job_title)
ds_salaries$job_title  <- factor(ds_salaries$job_title)

num_cols=ds_salaries[, c(2,11 )]
head(num_cols)
```

we encoded our ordinal and nominal variables using factor

## Removing irrelevant and duplicate data from the dataset

our dataset has 2 salary coulmns salary, and salary in usd, for this, we will measure the correlation between them

```{r}

cor(ds_salaries$salary,ds_salaries$salary_in_usd)
```

we see that salary and salary_in_usd are highly correlated so we will only take the salary

using chi-x we will measure the correlation between the variables employee residence and company_location

```{r}
chisq.test(ds_salaries$employee_residence,ds_salaries$company_location,correct = FALSE)
```

we reject the null hypothesis, x-squared\>\>\> p -value they are strongly correlated so we will only take company_location

```{r}
library(tidyverse)
ds_salaries <- ds_salaries %>%
  select(work_year,salary,experience_level,employment_type,job_title,salary_currency,company_location,company_size,remote_ratio)
  head(ds_salaries)
```

## 
