---
title: "R Notebook"
output: html_notebook
github_document: default
---
##Goal of collecting this Dataset 
Predicting Data Science job salaries using classification.

##The source of the dataset
Link to the source Dataset:
<https://www.kaggle.com/datasets/arnabchaki/data-science-salaries-2023?resource=download>

##Informition about the Dataset 
Number of Attributes: 11
Type of Attributes: nominal, ordinal, numeric
Number of objects: 3755
Class name: salary

  |Attributes name| Description| Data type |Possible values|
    |---------------|------------|-----------|---------------|
    |work_year      |The year the salary was paid| ordinal |Range between 2020-2023|
    |experience_level|The experience level in the job during the year| ordinal|(SE,MI,EN,EX)|
    |employment_type|The type of employment for the role| ordinal| (FT,PT,CT,FL)|
    |job_title| The role worked in during the year| nominal|(Data Engineer, Data Scientist,Data Analyst, Machine Learning Engineer, Analytics Engineer)|
    |salary|The total gross salary amount paid| numeric interval|Range Between [6000 - 30.4m]|
    |salary_currency|The currency of the salary paid as an ISO 4217 currency code|nominal|(USD,EUR,GBP,INR,CAD)|
    | salaryinusd        |  The salary in USD      |    numeric interval| Between[5132-450k]           |
    |employee_residence |Employee's primary country of residence in during the work year as an ISO 3166 country code| nominal |(US,GB,CA,ES,IN)|
    |remote_ratio |The overall amount of work done remotely |numeric Ratio | Range Between [0-100]|
    |company_location| The country of the employer's main office or contracting branch| nominal |(US,GB,CA,ES,IN)|
    |company_size | The median number of people that worked for the company during the year| ordinal |(S,M,L)|

##uploding the data
```{r}
ds_salaries= read.csv(url("https://raw.githubusercontent.com/DeemAlshaye/Miningproject/main/dataset/ds_salaries.csv"), header=TRUE)
View(ds_salaries)
```

```{r}
library(dplyr)
```

##summray the data
```{r}
str(ds_salaries)
```
##sample
```{r}

library(dplyr)
samp = sample_n(ds_salaries,10)
print(samp)
```

##missing data
```{r}
sum(is.na(ds_salaries))
```

##statical mesurment
```{r}
summary(ds_salaries$salary)
```

```{r}
summary(ds_salaries$salary_in_usd)
```

```{r}
summary(ds_salaries$remote_ratio)
```

```{r}
var(ds_salaries$salary)
var(ds_salaries$salary_in_usd)
var(ds_salaries$remote_ratio)
```


